---
title: "Code to calculate useful-stage EROIs for fossil fuels and the EROIs equivalent of renewable energy"
output: github_document
bibliography: References.bib
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```


## Introduction

This repository contains the code needed to replicate the analysis conducted in the Nature Energy paper "Estimation of useful-stage energy returns on investment for fossil fuels and implications for renewable energy systems" ([@aramendiaUsefulStageEROIs](https://doi.org/10.1038/s41560-024-01518-6)).

Besides the code contained in this repository, specific input data is required, see "Reproducing the analysis."

The analysis is structured using a [targets](https://books.ropensci.org/targets/) pipeline.


## Repository structure

The repository is structured as follows:

+ The "_targets.R" file contains the pipeline;
+ The "setup.R" file allows to define paths to datasets and to change the time period of analysis;
+ The "Functions" folder contains all the scripts defining the functions used in the "_targets.R" file;
+ The "Figures_Code" folder contains all the scripts required to reproduce the paper's figures once the pipeline has been calculated;
+ The "inst" file contains some required input data (like scenario data to conduct the intermittency analysis);
+ The "renv" folder contains files required to save the library environment using [renv](https://rstudio.github.io/renv/articles/renv.html).


## Reproducing the analysis

To run the analysis, the first action is to make sure that the required datasets have been correctly downloaded and that the "setup.R" script is correctly used to locate them. Specifically, the following datasets are required:

+ The [International Energy Agency](https://www.iea.org/) (IEA) Extended World Energy Balances (EWEB). An IEA license is required to access the dataset. Then, the path to the dataset needs to be indicated using the "iea_weeb_file_setup" variable in the "setup.R" file;
+ [Exiobase v3.8.2](https://zenodo.org/records/5589597): This dataset is free to access and download. The "IOT_YYYY_ixi.zip" files need to be downloaded for years 1995-2015 (the code takes care of unzipping the files). The path to the dataset needs to be indicated using the "path_to_exiobase_setup" variable in the "setup.R" file;
+ The Physical Supply Use Tables describing the Country-Level Primary-Final-Useful energy and exergy balances [@pfudatabase](https://doi.org/10.5518/1199). These balances include the raw EWEB IEA data and are therefore also subject to the IEA license. The dataset may be shared to any person having access to IEA data via institutional or personal license. The path to the dataset needs to be indicated using the "gpfu_psut_file_setup" variable in the "setup.R" file.

Note that when specifying in the "setup.R" the filepaths to datasets, it is expected that all datasets are in the same overarching folder indicated by the "path_to_datasets_setup" variable. So, if the IEA data is stored in the "Documents/Datasets/IEA/" folder, then "path_to_datasets_setup" should be set to "Documents/Datasets/" and "iea_weeb_file_setup" to "IEA/file.csv".

Then, on needs to load the library environment saved with [renv](https://rstudio.github.io/renv/articles/renv.html) by running:

```{r}
renv::restore()
```

To trigger the targets pipeline, one can then run the following line:

```{r}
targets::tar_make()
```


**Important caveat**: The pipeline is very heavy and running it on a standard desktop or laptop is highly likely to fail and to cause the machine to freeze. This code was run on a server with a RAM of 120 GB. 

<!-- For (slightly) lower computing requirements, the "aggregate" argument of the "aggregate_specified_regional_iea_data" function when calculating the specified_aggregated_regional_iea_data target may be set to TRUE (note that the results will then be different than those of the paper). -->


## Related works and used packages

This works builds on previous work developing a Physical Supply Use Table framework ([@Heun:2018](https://doi.org/10.1016/j.apenergy.2018.05.109)), and then developing this into a Multi-Regional Physical Supply Use Table framework using the IEA's EWEB ([@aramendiaDevelopingMultiRegionalPhysical2022](https://doi.org/10.1016/j.apenergy.2021.118413)).

Throughout this progress, a set of helpful packages, all of which are used in this pipeline, have been developed:

+ [IEATools](https://github.com/MatthewHeun/IEATools): allows to load, clean, and specify the IEA's EWEB, and then to construct Physical Supply Use Tables representing the IEA's EWEB;
+ [ECCTools](https://github.com/earamendia/ECCTools): provides functions to build Multi-Regional Physical Supply Use Tables;
+ [Recca](https://github.com/MatthewHeun/Recca): provides functions to perform input-output calculations using the Physical Supply Use Tables built from IEA data;
+ [EROITools](https://github.com/earamendia/EROITools): provides functions to aggregate the EROI calculations performed at the product level using the Recca package.


## References
